# Analysis of longitudial data

## RATS DATA
```{r}
library(tidyverse)

RATSL <- read.csv("RATSL.csv")


RATSL$ID <- factor(RATSL$ID)
RATSL$Group <- factor(RATSL$Group)


```

First part of this week's analysis investigates `r length(levels(RATSL$Group)) different groups of rats with different diets.


```{r}
library(ggplot2)

# Plot the RATSL data
ggplot(RATSL, aes(x = Time, y = Weight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  scale_y_continuous(name = "standardized bprs")
```
After looking at the plots, it is evident that gropu 1 rats are lower weight than other groups. Also, group 3 rats seem heavier than rats in group 2, except for an outlier rat in group 2. 

In all three groups, the rats have gained weight during the study (upward lines).

By standardizing the weight values, we may better analyze visually the weight gain pace of each group.
$$standardised(x) = \frac{x - mean(x)}{ sd(x)}$$

```{r}
# standardise RATSL data
RATSL <- RATSL %>% 
  group_by(Group) %>% 
  mutate(Weight_std = (Weight - mean(Weight))/sd(Weight)) %>% 
  ungroup()

# plot standardized weights
ggplot(RATSL, aes(x = Time, y = Weight_std, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  scale_y_continuous(name = "standardized bprs")
  
```
Weight gain seems faster in groups 1 & 3 (steeper line angle). Also, there seems to be one rather distinct outlier in each group.


```{r}

# Summary data with mean and standard error of bprs by treatment and week 
RATSLS <- RATSL %>%
  group_by(Group, Time) %>%
  summarise( mean = mean(Weight), se = sd(Weight)  ) %>%
  ungroup()

# Glimpse the data
glimpse(RATSLS)

# Plot the mean profiles
ggplot(RATSLS, aes(x = Time, y = mean, linetype = Group, shape = Group)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2,3)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2,3)) +
  geom_errorbar(aes(ymin=mean - se, ymax= mean + se, linetype="1"), width=0.3) +
  theme(legend.position = c(0.8,0.4)) +
  scale_y_continuous(name = "mean(Weight) +/- se(Weight)")

```
Group 3 is on average the heaviest. Also, it seems slightly that group 2 has the fastest weight gain on average.


```{r}
RATSL2 <- RATSL %>%
#  filter(Time > 0) %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()

# Glimpse the data
glimpse(RATSL2)

# Draw a boxplot of the mean versus treatment
ggplot(RATSL2, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", fill = "white") +
  scale_y_continuous(name = "mean(Weight)")

# Remove outliers
RATSL21 <- RATSL2 %>%
  filter(mean < 550 & mean > 240 )

ggplot(RATSL21, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", fill = "white") +
  scale_y_continuous(name = "mean(Weight)")
min(RATSL21$mean)
```
It seems that outlier especially in group 2 has an affect on the results.

```{r}
# Perform a two-sample t-test
# Make group lists
groups <- list()
for(i in 1:3) groups[[i]] <- RATSL21$mean[which(RATSL21$Group == i)]

# 1 vs. 2
t.test(groups[[1]], groups[[2]])

# 1 vs. 3
t.test(groups[[1]], groups[[3]])

# 2 vs. 3
t.test(groups[[2]], groups[[3]])
```
As notified earlier, there are clear differences between groups mean weights. Additionally, t-test reveals that the differences are statistically significant. 


## BPRS DATA
```{r}
# Load data

BPRSL <- read.csv("BPRSL.csv")

# Factor treatment & subject
BPRSL$treatment <- factor(BPRSL$treatment)
BPRSL$subject <- factor(BPRSL$subject)
# Convert to long form
# BPRSL <-  pivot_longer(BPRS, cols = -c(treatment, subject),
#                       names_to = "weeks", values_to = "bprs") %>%
#  arrange(weeks)

# Extract the week number
BPRSL <-  BPRSL %>% 
  mutate(week = as.integer(substr(weeks, 5, 5)))

# Plot data
ggplot(BPRSL, aes(x = week, y = bprs, linetype = subject)) +
  geom_line() +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme(legend.position = "none")
```

In both treatment groups there seems to me a general decreasing trend in BPRS. Subjects in the first group however seem more identical because there is less variance between subjects. 

```{r}
# Create linear regression model
BPRS_reg <- lm(bprs ~ week + treatment, data=BPRSL)

summary(BPRS_reg)

```

The regression model results show that there is a statistically significant relationship with variable week (expressing duration) and BPRS. Moreover, the decrease by week is on average ~2.3 points. However, no statistical significance between treatments. 


```{r}
library(lme4)

str(BPRSL)
BPRS_lme <- lmer(bprs ~ week + treatment + (1 | subject), data = BPRSL, REML = FALSE)
summary(BPRS_lme)
```


```{r}

# create a random intercept and random slope model

BPRS_lme2 <- lmer(bprs ~ week + treatment + (week | subject), data = BPRSL, REML = FALSE)
summary(BPRS_lme2)

```

No major differences between the two models.

```{r}

anova(BPRS_lme, BPRS_lme2)
```

```{r}
# Create a random intercept and slope model for a group × time interaction.
BPRS_lme3 <- lmer(bprs ~ week + treatment + (week | subject) + week * treatment, data = BPRSL, REML = FALSE)
summary(BPRS_lme3)

```
```{r}
# Create data with fitted valus

BPRSL_fit <- BPRSL
BPRSL_fit$bprs <- predict(BPRS_lme3)


ggplot(BPRSL_fit) +
  geom_line(aes(x = week, y = bprs, linetype = subject)) +
  scale_x_continuous(name = "Week") +
  scale_y_continuous(name = "BPRS") +
  theme(legend.position = "top") +
  facet_grid(. ~treatment, labeller=label_both) +
  theme(legend.position = "none")
```

Above is the plotted fitted model for both treatments.
