# Logistic regression

###1) R markdown is created
```{r}
date()
```

###2) Load dataset and examine column names and dimensions
```{r}
library(tidyverse)

stu <- read_csv("alc.csv")

dim(stu)

colnames(stu)
```

The data set is a joined from questionnaires targeted for two Portuguese schools. Its attributes include student grades, demographic, social and school related features) 


###3) Variable selection
```{r}
glimpse(stu)

stu2 <- select(stu, absences, G3, freetime, goout, high_use)
  
```
I assume that higher levels of absences, freetime and going out are linked with higher alcohol consumption. Maybe it will lead to lower grades (G3) as well.


###4) Variable exploration and analysis
```{r}

# initialize a plot of high_use and G3
g1 <- ggplot(stu2, aes(x = high_use, y = G3))

# define the plot as a boxplot and draw it
g1 + geom_boxplot() + ylab("grade")

# initialize a plot of absences and high_use
g2 <- ggplot(stu2, aes(x = high_use, y = absences))

# define the plot as a boxplot and draw it
g2 + geom_boxplot() + ylab("absences")

g3 <- ggplot(stu2, aes(x = high_use, y = freetime))
g3 + geom_boxplot() + ylab("freetime")

g4 <- ggplot(stu2, aes(x= high_use, y = goout))
g4 + geom_boxplot() + ylab("Go out")

```
I would say that my hypothses were right except for freetime which does not seem to have a large effect. 

###5) Logistic regression
```{r}
library(dplyr)
# create log. regr. model with glm()
m <- glm(high_use ~ absences + G3 + freetime + goout, data = stu2, family = "binomial")

# print out a summary of the model
summary(m)
# print out the coefficients of the model
coef(m)

```

